---
title: "Group 13 R Code"
output: pdf_document
---

## Setup ##
```{r Setup, warning = FALSE}
setwd("C:/NUS/Y2S1/BT2101/GitHub Group Project/bt2101_project")
data <- read.table("C:/NUS/Y2S1/BT2101/GitHub Group Project/bt2101_project/card.csv", sep=",", 
                   skip=2, header=FALSE)
header <- scan("C:/NUS/Y2S1/BT2101/GitHub Group Project/bt2101_project/card.csv", sep=",", 
               nlines=2, what=character()) 
colnames(data) <- header[26:50]
#removing ID column
data <- data[-1]
#renaming default_payment_next_month column to default
colnames(data)[24] <- 'default'
#renaming PAY_0 to PAY_1
colnames(data)[6] <- 'PAY_1'
```


## Checking Missing Data & Duplicate Data ##
```{r Checking Missing Data & Duplicate Data, warning=FALSE}
library(dplyr)
sum(is.na(data))
sum(duplicated(data))
dup <- data[duplicated(data) | duplicated(data, fromLast = TRUE),]
dup[order(dup$LIMIT_BAL,dup$AGE),]
```


## Changing Categorical Attributes to Factors ##
```{r Changing Catergorical Attributes to Factors, warning=FALSE}
data2 <- data
factors <- c('SEX','EDUCATION','MARRIAGE','PAY_1','PAY_2','PAY_3','PAY_4',
             'PAY_5','PAY_6','default')
data2[,factors] <- lapply(data2[,factors] , factor)
```


## Checking Correlation Between Quantitative Attributes ##
```{r Checking Correlation Between Quantitative Attributes, warning=FALSE} 
library(corrplot)
res <- cor(as.matrix(data2[,c('LIMIT_BAL', 'AGE', 'BILL_AMT1', 'BILL_AMT2',
                              'BILL_AMT3', 'BILL_AMT4','BILL_AMT5', 'BILL_AMT6',
                              'PAY_AMT1', 'PAY_AMT2', 'PAY_AMT3', 'PAY_AMT4', 
                              'PAY_AMT5', 'PAY_AMT6')]))
corrplot(res,method="circle", tl.cex=0.5) 
corrplot(res,method="number", number.cex= 14/ncol(data2), tl.cex=0.4)
```


## Checking for Anomalies ##
```{r Checking for Anomalies, warning=FALSE}
summary(data2)
#checking all the levels in PAY1-6
levels(data2$PAY_1)
levels(data2$PAY_2)
levels(data2$PAY_3)
levels(data2$PAY_4)
levels(data2$PAY_5)
levels(data2$PAY_6)
```


## Transforming Data ##
```{r Transforming Data, warning=FALSE}
data_t <- data2 
data_t$EDUCATION[data_t$EDUCATION %in% c('4','5','6')] <- '0'
data_t[,factors] <- lapply(data_t[,factors] , factor)
summary(data_t)
```


## Further Exploration of Data - Defaults ##
```{r Further Exploration of Data - Defaults, warning=FALSE}
library(ggthemes)
library(ggplot2)

#Looking at defaults
ggplot(data_t, aes(x=default, fill=default)) + geom_bar() + theme_excel()
sum(data_t$default == 1)/length(data$default)
```


## Further Exploration of Data - Limit Balance ##
```{r Further Exploration of Data - Limit Balance, warning=FALSE}
#Looking at LIMIT_BAL
ggplot(data_t, aes(x=LIMIT_BAL, fill = default), breaks=150) + 
  geom_histogram(position='dodge') + labs(x = 'limit balance') +
  theme_excel()
#percentage of clients with limit balance between 31250 and 62500 that defaulted
sum(data_t[31250<=data_t$LIMIT_BAL & data_t$LIMIT_BAL<=62500,]$default ==1)/ 
  length(data_t[31250<=data_t$LIMIT_BAL & data_t$LIMIT_BAL<=62500,]$default)
```


## Further Exploration of Data - Education ##
```{r Further Exploration of Data - Education, warning=FALSE}
#Looking at EDUCATION
ggplot(data_t, aes(x=EDUCATION, fill = default)) + geom_bar(position='dodge') + 
  geom_text(aes(label=..count..), stat="count", position =
              position_dodge(0.9), vjust=-0.15) + labs(x = 'education') + 
  theme_excel()
#count()#percentage of defaults in each category of Education
sum(data_t[data_t$EDUCATION == 0,]$default ==1) / length(data_t[data_t$EDUCATION 
                                                                == 0,]$default)
sum(data_t[data_t$EDUCATION == 1,]$default ==1) / length(data_t[data_t$EDUCATION
                                                                == 1,]$default)
sum(data_t[data_t$EDUCATION == 2,]$default ==1) / length(data_t[data_t$EDUCATION 
                                                                == 2,]$default)
sum(data_t[data_t$EDUCATION == 3,]$default ==1) / length(data_t[data_t$EDUCATION 
                                                                == 3,]$default)
```


## Further Exploration of Data - Marriage ##
```{r Further Exploration of Data - Marriage, warning=FALSE}
#Looking at Marriage
ggplot(data_t, aes(x=MARRIAGE, fill = default)) + geom_bar(position='dodge') +
  geom_text(aes(label=..count..), stat="count", position=position_dodge(0.9),
            vjust=-0.15) + labs(x = 'marriage') + theme_excel()
sum(data_t[data_t$MARRIAGE == 0,]$default ==1) / length(data_t[data_t$MARRIAGE 
                                                               == 0,]$default)
sum(data_t[data_t$MARRIAGE == 1,]$default ==1) / length(data_t[data_t$MARRIAGE 
                                                               == 1,]$default)
sum(data_t[data_t$MARRIAGE == 2,]$default ==1) / length(data_t[data_t$MARRIAGE 
                                                               == 2,]$default)
sum(data_t[data_t$MARRIAGE == 3,]$default ==1) / length(data_t[data_t$MARRIAGE 
                                                               == 3,]$default)
```


## Further Exploration of Data - Age ##
```{r Further Exploration of Data - Age, warning=FALSE}
#Looking at AGE
ggplot(data_t, aes(x=AGE, freq  = 2, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'Age')
```


## Further Exploration of Data - Payment Status ##
```{r Further Exploration of Data - Payment Status, warning=FALSE}
#looking at PAY1
ggplot(data_t, aes(x=PAY_1, fill = default)) + geom_bar(position='dodge') + 
  geom_text(aes(label=..count..), stat="count", position=position_dodge(0.9), 
            vjust=-0.15, size=2.5) + labs(x = 'September') + theme_excel()
#looking at PAY2
ggplot(data_t, aes(x=PAY_2, fill = default)) + geom_bar(position='dodge') +   
  geom_text(aes(label=..count..), stat="count", position=position_dodge(0.9),
            vjust=-0.15, size=2.5) + labs(x = 'August') + theme_excel()
#looking at PAY3
ggplot(data_t, aes(x=PAY_3, fill = default)) + geom_bar(position='dodge') + 
  geom_text(aes(label=..count..), stat="count", position=position_dodge(0.9),
            vjust=-0.15, size=2.5) + labs(x = 'July') + theme_excel()
#looking at PAY4
ggplot(data_t, aes(x=PAY_4, fill = default)) + geom_bar(position='dodge') + 
  geom_text(aes(label=..count..), stat="count", position=position_dodge(0.9), 
            vjust=-0.15, size=2.5) + labs(x = 'June') + theme_excel()
#looking at PAY5
ggplot(data_t, aes(x=PAY_5, fill = default)) + geom_bar(position='dodge') + 
  geom_text(aes(label=..count..), stat="count", position=position_dodge(0.9), 
            vjust=-0.15, size=2.5) + labs(x = 'May') + theme_excel()
#looking at PAY6
ggplot(data_t, aes(x=PAY_6, fill = default)) + geom_bar(position='dodge') + 
  geom_text(aes(label=..count..), stat="count", position=position_dodge(0.9),
            vjust=-0.15, size=2.5) + labs(x = 'April') + theme_excel()
```


## Further Exploration of Data - Bill Amount ##
```{r Further Exploration of Data - Bill Amount, warning=FALSE}
#looking at BILL_AMT1
ggplot(data_t, aes(x=BILL_AMT1, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'September') + theme_excel()
#looking at BILL_AMT2
ggplot(data_t, aes(x=BILL_AMT2, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'August') + theme_excel()
#looking at BILL_AMT3
ggplot(data_t, aes(x=BILL_AMT3, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'July') + theme_excel()
#looking at BILL_AMT4
ggplot(data_t, aes(x=BILL_AMT4, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'June') + theme_excel()
#looking at BILL_AMT5
ggplot(data_t, aes(x=BILL_AMT5, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'May') + theme_excel()
#looking at BILL_AMT6
ggplot(data_t, aes(x=BILL_AMT6, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'April') + theme_excel()
```

 
## Further Exploration of Data - Paid Amount ##
```{r Further Exploration of Data - Paid Amount, warning=FALSE}
#looking at PAY_AMT1
ggplot(data_t, aes(x=PAY_AMT1, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'September') + theme_excel()
#looking at PAY_AMT2
ggplot(data_t, aes(x=PAY_AMT2, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'August') + theme_excel()
#looking at PAY_AMT3
ggplot(data_t, aes(x=PAY_AMT3, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'July') + theme_excel()
#looking at PAY_AMT4
ggplot(data_t, aes(x=PAY_AMT4, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'June') + theme_excel()
#looking at PAY_AMT5
ggplot(data_t, aes(x=PAY_AMT5, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'May') + theme_excel()
#looking at PAY_AMT6
ggplot(data_t, aes(x=PAY_AMT6, fill = default)) + 
  geom_histogram(position='dodge') + labs(x = 'April') + theme_excel()
```


## Splitting into Test and Train Data ##
```{r Splitting into Test and Train Data, warning=FALSE}
set.seed(111)
n = nrow(data_t)
index <- 1:nrow(data_t)
testindex <- sample(index, trunc(n)/4) 
test.data <- data_t[testindex,] 
train.data <- data_t[-testindex,]
```


## Standardising Training Data ##
```{r Standardising Training Data, warning=FALSE}

data_s <- train.data
data_s$LIMIT_BAL<-scale(train.data$LIMIT_BAL)
data_s$AGE<-scale(train.data$AGE)
data_s$BILL_AMT1<-scale(train.data$BILL_AMT1)
data_s$BILL_AMT2<-scale(train.data$BILL_AMT2)
data_s$BILL_AMT3<-scale(train.data$BILL_AMT3)
data_s$BILL_AMT4<-scale(train.data$BILL_AMT4)
data_s$BILL_AMT5<-scale(train.data$BILL_AMT5)
data_s$BILL_AMT6<-scale(train.data$BILL_AMT6)
data_s$PAY_AMT1<-scale(train.data$PAY_AMT1)
data_s$PAY_AMT2<-scale(train.data$PAY_AMT2)
data_s$PAY_AMT3<-scale(train.data$PAY_AMT3)
data_s$PAY_AMT4<-scale(train.data$PAY_AMT4)
data_s$PAY_AMT5<-scale(train.data$PAY_AMT5)
data_s$PAY_AMT6<-scale(train.data$PAY_AMT6)
```


## PCA Transformation ##
```{r PCA Transformation & Analysis, warning=FALSE}
data_s_pca <- prcomp(data_s[,c(1,5,12:23)], center = TRUE, scale. = TRUE)
summary(data_s_pca)
str(data_s_pca)

library(ggfortify)
autoplot(data_s_pca, data = data_s, colour = 'default',
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 3)
```


## T.test Function ##
```{r T.test Function, warning=FALSE}
train.class <- train.data$default
tmatrix <- matrix(0,23,2) 
for (i in 1:23){ 
  x <- train.data[,i]
  results <- t.test(as.numeric(x) ~ train.class, var.equal= TRUE) 
  tmatrix[i,1] =results$statistic
  tmatrix[i,2] = results$p.value
}
df <- data.frame(tmatrix[,1],tmatrix[,2],names(data)[1:23]) 
names(df) <- c("t-value","p-value","attribute")
df
```


## Feature selection using Stepwise backward regression ##
```{r Feature selection using Stepwise backward regression, warning=FALSE}
library(leaps)
Outbackward <- regsubsets(train.class ~ ., data = train.data[, -24], 
                          method = "backward")
summary(Outbackward)
plot(Outbackward,scale="r2")
```


## Feature selection using Stepwise forward regression ##
```{r Feature selection using Stepwise forward regression, warning=FALSE}
library(leaps)
Outforward <- regsubsets(train.class ~ ., data = train.data[, -24],
                         method = "forward")
summary(Outforward)
summary(Outforward)
plot(Outforward,scale="r2")
```


## Chi-square statistic for Categorical Attribute ##
```{r Chi-square statistic for Categorical Attribute, warning=FALSE}
chistat <- matrix(0,23,2) 
issue <- rep(" ",23)
col <- ncol(train.data) - 1
class <- as.factor(train.data[,24])
for (I in 1: col) {
  x <- train.data[,I]
  tbl <- table(x,class)
  chi2res <- chisq.test(tbl) 
  chistat[I,1] <- chi2res$statistic
  chistat[I,2] <- chi2res$p.value 
}
df <- data.frame(chistat[,1:2],names(data)[1:23]) 
names(df) <- c("chi2 stat","p-value","issue") 
df
df[order(df[,1]),]
```


## Decision Tree - CART ##
```{r Decision Tree - CART, warning=FALSE}
library(partykit)
library(rpart)
library(rpart.plot)
cart.tree1 <- rpart(class~., data=train.data[names(data)[1:23]], model=TRUE, 
                    method='class', control = rpart.control(cp = 0.001))
cart1.fit <- predict(cart.tree1,data=train.data[names(data)[1:23]],type="class") 
table(cart1.fit,class)
rpart.plot(cart.tree1)
summary(cart.tree1)
```


## Decision Tree - C5 ##
```{r Decision Tree - C5, warning=FALSE}
library(C50)
C50tree <- C5.0( class ~ .,data= train.data[names(data)[1:23]])
summary(C50tree)

```


## Load libraries ##
```{r Load libraries, warning=FALSE}
library(caret)
library(ipred)
library(randomForest)
library(gbm)
library(InformationValue)
library(naivebayes)
library(nnet)
library(e1071)
library(pROC)

```


## Creating the formula for Modelling ##
```{r Creating the formula for Modelling, warning=FALSE}
formula <- as.formula(default ~ PAY_1 + PAY_3)
formula

```


## chi-square test to check correlation ##
```{r chi-square test to check correlation, warning=FALSE}
table1 <- table(train.data$PAY_1, train.data$PAY_3)
chisq.test(table1)

```


## Ensemble Method - Bagging ##
```{r Ensemble Method - Bagging, warning=FALSE}

set.seed(111)
bagging <- bagging(formula, data = train.data, nbagg = 50, coob = TRUE)

fittedTrain <- predict(bagging, data = train.data[c("PAY_1","PAY_3")])
fittedTrainProb <- predict(bagging, data = train.data[c("PAY_1","PAY_3")], 
                           type = "prob")
actualTrain <- train.data$default
trainMatrix <- caret::confusionMatrix(fittedTrain, actualTrain, positive = '1')
trainMatrix

fittedTest <- predict(bagging, newdata = test.data[c("PAY_1","PAY_3")])
fittedTestProb <- predict(bagging, newdata = test.data[c("PAY_1","PAY_3")], 
                          type = 'prob')
actualTest <- test.data$default
testMatrix <- caret::confusionMatrix(fittedTest, actualTest, positive = '1')
testMatrix

#ROC#

training.probs <- predict(bagging, newdata = train.data[c("PAY_1","PAY_3")], 
                          type = "prob")
plotROC(actuals=actualTrain,predictedScores=training.probs[,2])

test.probs <- predict(bagging, newdata = test.data[c("PAY_1","PAY_3")], 
                      type = "prob")
plotROC(actuals=actualTest,predictedScores=test.probs[,2])

```


## Ensemble Method - Random Forest ##
```{r Ensemble Method - Random Forest, warning=FALSE}

set.seed(111)
rf <- randomForest(formula, data = train.data)

fittedTrain <- predict(rf, data = train.data[c("PAY_1","PAY_3")]) 
actualTrain <- train.data$default
trainMatrix <- caret::confusionMatrix(fittedTrain, actualTrain, positive = '1')
trainMatrix


fittedTest <- predict(rf, newdata = test.data[c("PAY_1","PAY_3")]) 
actualTest <- test.data$default
testMatrix <- caret::confusionMatrix(fittedTest, actualTest, positive = '1')
testMatrix


#ROC#

fittedTrainProb <- predict(rf, data = train.data[c("PAY_1","PAY_3")], 
                           type = "prob") 
plotROC(actuals = actualTrain, predictedScores = fittedTrainProb[,2])

fittedTestProb <- predict(rf, newdata = test.data[c("PAY_1","PAY_3")], 
                          type = "prob")
plotROC(actuals = actualTest, predictedScores = fittedTestProb[,2])

```

 
## Naive Bayes ##
```{r Naive Bayes, warning=FALSE}

set.seed(111)
nb <- naive_bayes(formula , data = train.data, laplace = 0.1)

fittedTrain <- predict(nb, data = train.data[c("PAY_1","PAY_3")], 
                       type = "class")
actualTrain <- train.data$default
trainMatrix <- caret::confusionMatrix(fittedTrain, actualTrain, positive = '1')
trainMatrix

fittedTest <- predict(nb, newdata = test.data[c("PAY_1","PAY_3")], 
                      type = "class")
actualTest <- test.data$default
testMatrix <- caret::confusionMatrix(fittedTest, actualTest, positive = '1')
testMatrix

#ROC#

fittedTrain1 <- predict(nb, data = train.data[c("PAY_1","PAY_3")], 
                        type = "prob")
fittedTest1 <- predict(nb, newdata = test.data[c("PAY_1","PAY_3")], 
                       type = "prob")

plotROC(actuals = actualTrain, predictedScores = fittedTrain1[,2])
plotROC(actuals = actualTest, predictedScores = fittedTest1[,2])

```


## Logistic Regression ##
```{r Logistic Regression, warning=FALSE}

set.seed(111)
logReg <- glm(formula , data = train.data, family = "binomial")

train.pred <- predict(logReg, data = train.data[c("PAY_1","PAY_3")],
                      type = "response")
actualTrain <- train.data$default
optCutTrain <- optimalCutoff(actualTrain, train.pred, 
                             optimiseFor="misclasserror")
fittedTrain <- as.factor(ifelse(train.pred < optCutTrain, 0, 1))
trainMatrix <- caret::confusionMatrix(fittedTrain, actualTrain, positive = '1')
trainMatrix

test.pred <- predict(logReg, newdata = test.data[c("PAY_1","PAY_3")], 
                     type = "response")
actualTest <- test.data$default
optCutTest <- optimalCutoff(actualTest, test.pred, 
                             optimiseFor="misclasserror")
fittedTest <- as.factor(ifelse(test.pred < optCutTest, 0, 1))
testMatrix <- caret::confusionMatrix(fittedTest, actualTest, positive = '1')
testMatrix

#ROC#

#train set
plotROC(actuals = actualTrain, predictedScores = train.pred)

#test set
plotROC(actuals = actualTest, predictedScores = test.pred)


```


## SVM ##
```{r SVM, warning=FALSE}

set.seed(111)
svm.model <- svm(formula, data = train.data, type = "C-classification", 
                 kernel = "linear")

fittedTrain <- predict(svm.model, data = train.data[c("PAY_1","PAY_3")])
actualTrain <- train.data$default
trainMatrix <- caret::confusionMatrix(fittedTrain, actualTrain, positive = '1')
trainMatrix

fittedTest <- predict(svm.model, newdata = test.data[c("PAY_1","PAY_3")])
actualTest <- test.data$default
testMatrix <- caret::confusionMatrix(fittedTest, actualTest, positive = '1')
testMatrix

#ROC#

fittedTrain1 <- predict(svm.model, newdata = train.data[c("PAY_1","PAY_3")], 
                        type = "prob")
plotROC(actuals=actualTrain,predictedScores=as.numeric(fittedTrain1))

fittedTest1 <- predict(svm.model, newdata = test.data[c("PAY_1","PAY_3")], 
                       type = "prob")
plotROC(actuals=actualTest,predictedScores=as.numeric(fittedTest1))

```










